---
title: "R Notebook"
output: html_notebook
---

#Introduction
The goal here is to use the biological replicates of the N2, N. parisii, and N. ausubeli experiments to attempt to understand the biological variability of gene counts. I will do this by comparing biological replicates against themselves, N2 sample 1 compared to N2 sample 2 for example, in order to develop lists of genes that are "differentially experssed" within biological replicates. Ultimately, the goal is to develope a qualitative set of filters that we can use to identify genes of interest in the experiments that make use of single replicates.

#Step 1 - load data
```{r}
files2 <- read.csv("nematocida.counts.matrix", stringsAsFactors = F, check.names = F)
count_matrix <- as.matrix(files2[,-1])
rownames(count_matrix) <- files2[,1]
head(count_matrix)
x3 <- DGEList(counts = count_matrix)
class(x3)
head(x3$counts)
x3$samples
```

#Step 2 - organize data
Normally, data would be grouped by experimental factors, ie strain or treatment or temperature, here however, I am placing each sample into its own individual group.
```{r}
group <- as.factor(1:12)
x3$samples$group <- group 
dim(x3)
x3$samples
```

#Step 3 - data pre-processing
Remove genes with low counts. TMM normalization
```{r}
cpm3 <- cpm(x3)
keep.exprs <- rowSums(cpm3>1)>=1
x3 <- x3[keep.exprs,, keep.lib.sizes=F]

dim(x3)
x3 <- calcNormFactors(x3, method = "TMM")
x3$samples$norm.factors

bcv <- 0.02
```

#Step 4 - experiment design and contrast matrix construction
I am skipping several steps here, including unsuperivised clustering analysis.
```{r eval=F}
design.matrix <- model.matrix(~0+group)
colnames(design.matrix) <- substring(colnames(design.matrix), 12, nchar(colnames(design.matrix)))
colnames(design.matrix) <- gsub("-","",colnames(design.matrix))
contrast.matrix <- makeContrasts(N2.1 = N2_S1_L004 - N2_S6_L004,
                                 N2.2 = N2_S1_L004 - N2_S15_L004,
                                 N2.3 = N2_S6_L004 - N2_S15_L004,
                                 np.1 = N2ERTm1_S4_L004 - N2ERTm1_S9_L004,
                                 np.2 = N2ERTm1_S4_L004 - N2ERTm1_S18_L004,
                                 np.3 = N2ERTm1_S9_L004 - N2ERTm1_S18_L004,
                                 na.1 = N2ERTm2_S5_L004 - N2ERTm2_S10_L004,
                                 na.2 = N2ERTm2_S5_L004 - N2ERTm2_S19_L004,
                                 na.3 = N2ERTm2_S10_L004 - N2ERTm2_S19_L004, levels=design.matrix)
```
#Perform pairwise comparisons
I am comparing the different conditions using an artificial dispersion coefficient. The p-values and FDR values are nonsensical. IGNORE THEM!
```{r}
et <- exactTest(x3, pair = c(1,4), dispersion = bcv^2)
topTags(et, n=Inf, sort.by = "logFC")
dfnew2$mock1 <- et$table$logFC
```

```{r}
et <- exactTest(x3, pair = c(1,7), dispersion = bcv^2)
topTags(et, n=Inf, sort.by = "logFC")
dfnew2$mock2 <- et$table$logFC
```

```{r}
et <- exactTest(x3, pair = c(4,7), dispersion = bcv^2)
topTags(et, n=Inf, sort.by = "logFC")
dfnew2$mock3 <- et$table$logFC
```

```{r}
et <- exactTest(x3, pair = c(2,5), dispersion = bcv^2)
topTags(et, n=Inf, sort.by = "logFC")
dfnew2$np1 <- et$table$logFC
```

```{r}
et <- exactTest(x3, pair = c(2,8), dispersion = bcv^2)
topTags(et, n=Inf, sort.by = "logFC")
dfnew2$np2 <- et$table$logFC
```

```{r}
et <- exactTest(x3, pair = c(5,8), dispersion = bcv^2)
topTags(et, n=Inf, sort.by = "logFC")
dfnew2$np3 <- et$table$logFC
```

```{r}
et <- exactTest(x3, pair = c(3,6), dispersion = bcv^2)
topTags(et, n=Inf, sort.by = "logFC")
dfnew2$nd1 <- et$table$logFC
```

```{r}
et <- exactTest(x3, pair = c(3,9), dispersion = bcv^2)
topTags(et, n=Inf, sort.by = "logFC")
dfnew2$nd2 <- et$table$logFC
```

```{r}
et <- exactTest(x3, pair = c(6,9), dispersion = bcv^2)
topTags(et, n=Inf, sort.by = "logFC")
dfnew2$nd3 <- et$table$logFC
```

```{r}
head(dfnew2)
write.csv(dfnew2, file = "2018_07_25_replicate_logFC.csv", quote = F)
```

